---
title: "01-Preprocessing"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "01-Preprocessing" # change if you rename file
---

```{r here, message=FALSE}
here::i_am(paste0(params$name, ".Rmd"), uuid = "664c6b80-49a5-449b-ae40-2ff6c9851b54")
```

The purpose of this document is ...

```{r packages}
library(tidyverse)
library(flowCore)
library(flowClust)
library(ggcyto)
library(gridExtra)
library(knitr)
library(ggplot2)
library(flowAI)
```

```{r directories}
# create or *empty* the target directory, used to write this file's data: 
projthis::proj_create_dir_target(params$name, clean = TRUE)

# function to get path to target directory: path_target("sample.csv")
path_target <- projthis::proj_path_target(params$name)

# function to get path to previous data: path_source("00-import", "sample.csv")
path_source <- projthis::proj_path_source(params$name)

```

## Import Data

```{r import, eval=FALSE}
fcs.direction <- "~/FCS/Z-Project Bacillus_pretest/fcs"
z_fs <- read.flowSet(path = fcs.direction,alter.names = T,transformation = F)
z_gate <- read.FCS("~/FCS/Z-Project Bacillus_pretest/Gating.fcs",alter.names = T,transformation=F)
markernames(z_fs[[1]])
```

### Basic Visualization

We select FSC, which represents the cell size and the fluorescence with a wavelength of 460 nm to generate the raw scatter plot.

```{r visualization, message=FALSE}
library(ggplot2)
ggplot(z_fs[[1]],aes(x=`PMT.1`,y=`PMT.9`))+
  geom_hex(bins=100)+
  theme_bw()+
  labs(x="FSC",y="Fluorescence",title = "FSC vs Fluoresence",subtitle = "Inner_zone_DAPI")+
  xlim(0,2500)+ylim(0,2500)+coord_fixed()
```

## Compensation + Transformation

In this section, we take one flowFrame, the Inner_Zone_DAPI, of whole flowSet as an example to illustrate the effect of compensation and transformation.

```{r comp,eval=FALSE}
# Compensation
y_comp <- compensate(z_fs[[1]],spillover(z_fs[[1]])$SPILL)

```

```{r trans}
# Transformation using log-transformation
trans <- estimateLogicle(y_comp,colnames(y_comp[,15:42]))
y_comp_trans <- transform(y_comp,trans)

ggplot(y_comp_trans,aes(x=`PMT.1`,y=`PMT.9`))+
  geom_hex(bins=100)+
  theme_bw()+
  labs(x="FSC",y="Fluo.",title = "FSC vs Fluo. after log-transformation",subtitle = "Inner_zone_DAPI")

```

## Quality Assessment

```{r quality-check}
library(flowAI)
y_comp_clean <- flow_auto_qc(y_comp)
```

## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
