---
title: "02b-flowmix on microbial community data"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "02b-flowmix on microbial community data" # change if you rename file
---

```{r here, message=FALSE}
here::i_am(paste0(params$name, ".Rmd"), uuid = "989c1b7b-94e7-44a3-ba84-4dfb55c493e8")
```

The purpose of this document is ...

```{r packages,message=F}
library("conflicted")
library(readxl)
library(purrr)
library(dplyr)
library(flowmix)
library(flowCore)
library(flowWorkspace)
library(ggcyto)
library(tidyverse)
library(RColorBrewer)
library(flowEMMi)
```

```{r directories}
# create or *empty* the target directory, used to write this file's data: 
projthis::proj_create_dir_target(params$name, clean = TRUE)

# function to get path to target directory: path_target("sample.csv")
path_target <- projthis::proj_path_target(params$name)

# function to get path to previous data: path_source("00-import", "sample.csv")
path_source <- projthis::proj_path_source(params$name)
```

## Task1

In this part, we're focusing on microbial communities from paper: Neutral mechanisms and niche differentiation in steady-state insular microbial communities reveald by single cell analysis by Liu et al., 2019

This flow cytometry data consists of 5 reactors data.

C1 and C2 are two control reactors in order to test for neutral behaviour under undisturbed communities.

D1, D2 and D3 were disturbed by a repeated and soft temperature stressor from 30 to 40 degrees.

Liu et. al sampled from all five reactors and analyzed them by flow cytometry. All raw data can be accessed by flowRepository (<https://flowrepository.org/>) under accession number: FR-FCM-ZYWX.

### Import data

```{r import-data, message=F}
path_Liu <- "~/Desktop/FlowRepository_Liu/"
file_name_Liu <- list.files(path_Liu,pattern = "\\.fcs$",full.names = T)
C1_data <- list()

for (i in file_name_Liu){
  data <- read.FCS(i,alter.names = T,transformation = F)
  new_name <- substr(i,start = 43,stop=57)
  C1_data[[new_name]] <- data
}

```

```{r,echo=F}
lengthc1 <- length(C1_data)
print(paste("The whole experiment has",lengthc1,"time points"))
```

```{r}
test1 <-C1_data[[1]]
ggcyto(test1, aes(x = "PMT.1", y = "PMT.9")) + geom_hex(bins = 128)
```

Below is another visulization method supported by flowEMMI v2 package
```{r,echo=T,eval=F}
plotDensityAndEllipses(fcsData = test1, ch1="PMT.1", ch2="PMT.9",logScale = F,
                       axis_size=10, axisLabeling_size=10,
                       xlab = "Forward Scatter", ylab = "DAPI", font = "Arial")
```

### flowEMMI v2
Since each time point contains around 200k cells, every run takes about 30mins, here is an example of one time point.
```{r test}
fdo_test1 <- mkFlowDataObject(test1, xChannel="PMT.1", yChannel="PMT.9")

set.seed(1)
gating_test1 <- flowEMMi( fdo=fdo_test1
                       , xMin=1, xMax=60000, yMin=1, yMax=60000
                       , initFraction=0.01
                       , finalFraction=1.0
                       , minClusters=5, maxClusters=15, clusterbracket=2
                       , numberOfInits=5
                       , verbose=TRUE
                       , parallel=FALSE
                       , convergenceEpsilon=0.01
                       , whenToRemoveOverlaps = 20
                       , mergeWhenCenter = FALSE
                       , mergeWhenTwoCenters = FALSE
                       , thresholdForDeletion = 0.2
                       , threshold = 0.9
                       , considerWeights=TRUE
                       , plot = FALSE
                       , alpha=0.9
                       , minMinor=500
)

pic <- plotDensityAndEllipses(fcsData = test1, ch1="PMT.1", ch2="PMT.9", alpha=0.9,
                                 logScale = F, results = gating_test1$best,
                                 ellipseDotSize = 0.5, axis_size=10, axisLabeling_size=10,
                                 xlab = "Forward Scatter", ylab = "DAPI", font = "Arial")

```

### flowmix
First, all fcs datas need to be converted to the flowmix required format for downstream analysis.

FSC and DAPI fluorescence would be the features for the cytograms, and environmental variables come from the abiotic sheet.
```{r convert}
# cytograms
C1_ylist <- list()

for (i in seq_along(C1_data)){
  C1_ylist[[i]] <- C1_data[[i]]@exprs[,c(11,13,27)] # PMT.1/2 for FSC/SSC, PMT.9 for fluorescence
}

# environment variables
abio_data <- read_excel("~/Desktop/osfstorage/Data sheet for abiotic parameters.xlsx")
C1_X <- abio_data[abio_data$Reactor=="C1",4:11]
C1_X[is.na(C1_X)] <- 0
C1_X <- as.matrix(C1_X)
```

Notation: since in the original data, each time point contains over 200k cells, here I select 10k for each as subset for illustration. The results for whole dataset comes later.

```{r binning}
# binning data
set.seed(0)
ysub <- list()

for (i in 1:65) {
  rrow <- sample(1:200000,size=2000,replace = F)
  ysub[[i]] <- C1_ylist[[i]][rrow,]
  ysub[[i]][,] <- as.double(ysub[[i]][,])
}

grid <- make_grid(ysub,gridsize = 30)
obj <- bin_many_cytograms(ysub,grid,mc.cores = 4,verbose = F)
sub_ylist <- obj$ybin_list
sub_countslist <- obj$counts_list
```

### on-going
Havn't got the results for flowmix, but all required data have been prepared
```{r flowmix,eval=F}
# flowmix for 2d
set.seed(0)
sub_rest <- flowmix(sub_ylist,C1_X,
                   countslist=sub_countslist,numclust=5,
                   mean_lambda=0.001,prob_lambda=0.001,
                   verbose=F,maxdev=0.5)

```

## Task2

```{r import,message=F}
folder_path <- "~/Desktop/FlowRepo/"
file_name <- list.files(folder_path,pattern = "\\.fcs$",full.names = T)
MoFloA_Feb_data <- list()
MoFloA_Mar_data <- list()
MoFloA_Nov_data <- list()
MoFloA_Dec_data <- list()

# Group files by the month

for (file_name in file_name){
  month <- substr(file_name,start = 46,stop = 48) # use nchar to locate
  new_name <- substr(file_name,start = 43, stop = 58)
  data <- read.FCS(file_name,alter.names = T,transformation=F)
  if (month=="Feb") { MoFloA_Feb_data[[new_name]] <- data}
  else if (month=="Mar") {MoFloA_Mar_data[[new_name]] <- data}
  else if (month=="Nov") {MoFloA_Nov_data[[new_name]] <- data}
  else if (month=="Dec") {MoFloA_Dec_data[[new_name]] <- data}
}

```

### flowEMMI v2

```{r}
test <- MoFloA_Nov_data[[1]]

# FL4 is the signal channel
fdo_test <- mkFlowDataObject(test, xChannel="FS.Log", yChannel="FL.4.Log")

set.seed(1)
gating_test <- flowEMMi( fdo=fdo_test
                       , xMin=1, xMax=5000, yMin=1, yMax=5000
                       , initFraction=0.01
                       , finalFraction=1.0
                       , minClusters=5, maxClusters=15, clusterbracket=2
                       , numberOfInits=5
                       , verbose=TRUE
                       , parallel=FALSE
                       , convergenceEpsilon=0.01
                       , whenToRemoveOverlaps = 20
                       , mergeWhenCenter = FALSE
                       , mergeWhenTwoCenters = FALSE
                       , thresholdForDeletion = 0.2
                       , threshold = 0.9
                       , considerWeights=TRUE
                       , plot = FALSE
                       , alpha=0.9
                       , minMinor=500
)

plot_g <- plotDensityAndEllipses(fcsData = test, ch1="FS.Log", ch2="FL.4.Log", alpha=0.9,
                                 logScale = T, results = gating_test$best,
                                 ellipseDotSize = 0.5, axis_size=10, axisLabeling_size=10,
                                 xlab = "Forward Scatter", ylab = "DAPI", font = "Arial")

```

## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
