---
title: "02-Gating"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "02-Gating" # change if you rename file
---

```{r here, message=FALSE}
here::i_am(paste0(params$name, ".Rmd"), uuid = "3dc33fb3-330d-46b4-b31d-8d7355d8d04e")
```

The purpose of this document is ...

```{r packages,message=F}
library(flowEMMi)
library(tidyverse)
library(flowCore)
library(flowClust)
library(ggcyto)
library(gridExtra)
library(knitr)
library(ggplot2)
library(gridGraphics)
```

```{r directories}
# create or *empty* the target directory, used to write this file's data: 
projthis::proj_create_dir_target(params$name, clean = TRUE)

# function to get path to target directory: path_target("sample.csv")
path_target <- projthis::proj_path_target(params$name)

# function to get path to previous data: path_source("00-import", "sample.csv")
path_source <- projthis::proj_path_source(params$name)
```

# Cell Population Identification

The process to identify the specific cell subsets is known as gating.The gating methods can be divided in two main categories: sequential (Manual) Gating and Automated Gating. Automated gating is performed in two ways: unsupervised and supervised. In below we'll focus on two unsupervised approaches, flowmix and flowEMMI v2

## flowEMMI v2
```{r flowEMMI}
fcs.direction <- "~/FCS/Z-Project Bacillus_pretest/fcs"
z_fs <- read.flowSet(path = fcs.direction,alter.names = T,transformation = F)

location <- c("Inner_zone","Middle_zone","Outer_zone","Surrounding","Whole_colony")
plot_list <- list()

for (i in 1:5){
  data_name <- paste0(location[i],"_filter_DAPI.fcs")
  data <- z_fs[[data_name]]
  pic <- plotDensityAndEllipses(fcsData = data, ch1="PMT.1", ch2="PMT.9",
                            logScale = F, title = data_name,
                            axis_size=10, axisLabeling_size=10,
                            xlab = "Forward Scatter", ylab = "DAPI Fluorescence", 
                            font = "Arial")
  plot_list[[i]] <- pic$plot
}

grid.arrange(grobs = plot_list, ncol = 3)

```
```{r EM-algorithm,message=F}
set.seed(1)
gating_list <- list()
gating_plot_list <- list()

for (i in 1:5){
  data_name <- paste0(location[i],"_filter_DAPI.fcs")
  data <- z_fs[[data_name]]
  fdo <- mkFlowDataObject(data, xChannel="PMT.1", yChannel="PMT.9")
  gating_fcsData <- flowEMMi( fdo=fdo
                             , xMin=1500, xMax=50000, yMin=1500, yMax=50000
                             , initFraction=0.01
                             , finalFraction=1.0
                             , minClusters=5, maxClusters=15, clusterbracket=2
                             , numberOfInits=5
                             , verbose=TRUE
                             , parallel=FALSE
                             , convergenceEpsilon=0.01
                             , whenToRemoveOverlaps = 20
                             , mergeWhenCenter = FALSE
                             , mergeWhenTwoCenters = FALSE
                             , thresholdForDeletion = 0.2
                             , threshold = 0.9
                             , considerWeights=TRUE
                             , plot = FALSE
                             , alpha=0.9
                             , minMinor=500)
  gating_list[[i]] <- gating_fcsData$best
}

for (i in 1:5){
  data_name <- paste0(location[i],"_filter_DAPI.fcs")
  data <- z_fs[[data_name]]
  gating_pic <- plotDensityAndEllipses(fcsData = data, ch1="PMT.1", ch2="PMT.9",
                                       alpha=0.9,logScale = T, 
                                       results = gating_list[[i]],
                                       title = data_name,
                            ellipseDotSize = 0.5, axis_size=10, axisLabeling_size=10,
                            xlab = "Forward Scatter", ylab = "DAPI Fluorescence", 
                            font = "Arial")
  gating_plot_list[[i]] <- gating_pic$plot
}

```

## Clustering by robust model-based method
```{r flowClust}
library(flowClust)

for (i in 1:5){
  data_name <- paste0(location[i],"_DAPI.fcs")
  data <- z_fs[[data_name]]
  data_gate <- flowClust(data,varNames = c("PMT.1","PMT.2"), K=1,B=100)
  data_new <- data[data %in% data_gate,]
  data_gate_new <- flowClust(data_new,varNames = c("PMT.1","PMT.9"), 
                             K=1:5,B=100)
  criterion(data_gate_new,"BIC")
  ruleOutliers(data_gate_new[[5]]) <- list(level=0.95)
  plot(data_gate_new,data=data_new,level=0.8)
}
```



## Files written

These files have been written to the target directory, `r paste0("data/", params$name)`:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
